---
title: "Travel Time"
author: "Yifeng Fan"
date: "2022-12-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(esquisse)
library(ggplot2)
```

```{r}
#Import the railway dataset
hr_q4_21<- read.csv("HRTravelTimesQ4_21.csv")

lr_q4_21<- read.csv("LRTravelTimesQ4_21.csv")

hr_q1_22<- read.csv("2022-Q1_HRTravelTimes.csv")
hr_q2_22<- read.csv("2022-Q2_HRTravelTimes.csv")
hr_q3_22<- read.csv("2022-Q3_HRTravelTimes.csv")

lr_q1_22<- read.csv("2022-Q1_HRTravelTimes.csv")
lr_q2_22<- read.csv("2022-Q2_HRTravelTimes.csv")
lr_q3_22<- read.csv("2022-Q3_HRTravelTimes.csv")

```

```{r}
#Import the bus data
bs_11_21<- read.csv("MBTA-Bus-Arrival-Departure-Times_2021-11.csv")
bs_12_21<- read.csv("MBTA-Bus-Arrival-Departure-Times_2021-12.csv")
bs_01_22<- read.csv("MBTA-Bus-Arrival-Departure-Times_2022-01.csv")
bs_02_22<- read.csv("MBTA-Bus-Arrival-Departure-Times_2022-02.csv")
bs_03_22<- read.csv("MBTA-Bus-Arrival-Departure-Times_2022-03.csv")
bs_04_22<- read.csv("MBTA-Bus-Arrival-Departure-Times_2022-04.csv")
bs_05_22<- read.csv("MBTA-Bus-Arrival-Departure-Times_2022-05.csv")
bs_06_22<- read.csv("MBTA-Bus-Arrival-Departure-Times_2022-06.csv")
bs_07_22<- read.csv("MBTA-Bus-Arrival-Departure-Times_2022-07.csv")
bs_08_22<- read.csv("MBTA-Bus-Arrival-Departure-Times_2022-08.csv")
bs_09_22<- read.csv("MBTA-Bus-Arrival-Departure-Times_2022-09.csv")
bs_10_22<- read.csv("MBTA-Bus-Arrival-Departure-Times_2022-10.csv")

```


```{r}
#Pick the first week from each of the 12 months
#2021 Q4
hr_q4_21$week_num <- strftime(hr_q4_21$service_date, format = "%V")
#first week of each month: week 40, 44, 49 
hr_q4_week<- hr_q4_21 %>% filter(week_num %in% c(40, 44, 49) )


#2022 Q1
hr_q1_22$week_num <- strftime(hr_q1_22$service_date, format = "%V")
#first week of each month: week 01, 06, 10 
hr_q1_week<- hr_q1_22 %>% filter(week_num %in% c(01, 06, 10) )


#2022 Q2
hr_q2_22$week_num <- strftime(hr_q2_22$service_date, format = "%V")
#first week of each month: week 14, 18, 23 
hr_q2_week<- hr_q2_22 %>% filter(week_num %in% c(14, 18, 23) )


#2022 Q3
hr_q3_22$week_num <- strftime(hr_q3_22$service_date, format = "%V")
#first week of each month: week 27，31，36 
hr_q3_week<- hr_q3_22 %>% filter(week_num %in% c(27, 31, 36) )


#combine the heavy rail data
hr_week<- rbind(hr_q4_week, hr_q1_week, hr_q2_week, hr_q3_week)

```

```{r}
#Pick the first week from each of the 12 months
#2021 Q4
lr_q4_21$week_num <- strftime(lr_q4_21$service_date, format = "%V")
#first week of each month: week 40, 44, 49 
lr_q4_week<- lr_q4_21 %>% filter(week_num %in% c(40, 44, 49) )


#2022 Q1
lr_q1_22$week_num <- strftime(lr_q1_22$service_date, format = "%V")
#first week of each month: week 01, 06, 10 
lr_q1_week<- lr_q1_22 %>% filter(week_num %in% c(01, 06, 10) )


#2022 Q2
lr_q2_22$week_num <- strftime(lr_q2_22$service_date, format = "%V")
#first week of each month: week 14, 18, 23 
lr_q2_week<- lr_q2_22 %>% filter(week_num %in% c(14, 18, 23) )


#2022 Q3
lr_q3_22$week_num <- strftime(lr_q3_22$service_date, format = "%V")
#first week of each month: week 27，31，36 
lr_q3_week<- lr_q3_22 %>% filter(week_num %in% c(27, 31, 36) )


#combine the heavy rail data
lr_week<- rbind(lr_q4_week, lr_q1_week, lr_q2_week, lr_q3_week)

```

```{r}
#Pick 12 weeks from the bus dataset 

bs_11_21$week_num <- strftime(bs_11_21$service_date, format = "%V")
bs_11_21_week<- bs_11_21 %>% filter(week_num == "44" )

bs_12_21$week_num <- strftime(bs_12_21$service_date, format = "%V")
bs_12_21_week<- bs_12_21 %>% filter(week_num == "49" )

bs_01_22$week_num <- strftime(bs_01_22$service_date, format = "%V")
bs_01_22_week<- bs_01_22 %>% filter(week_num == "01" )

bs_02_22$week_num <- strftime(bs_02_22$service_date, format = "%V")
bs_02_22_week<- bs_02_22 %>% filter(week_num == "06" )

bs_03_22$week_num <- strftime(bs_03_22$service_date, format = "%V")
bs_03_22_week<- bs_03_22 %>% filter(week_num == "10" )

bs_04_22$week_num <- strftime(bs_04_22$service_date, format = "%V")
bs_04_22_week<- bs_04_22 %>% filter(week_num == "14" )

bs_05_22$week_num <- strftime(bs_05_22$service_date, format = "%V")
bs_05_22_week<- bs_05_22 %>% filter(week_num == "18" )

bs_06_22$week_num <- strftime(bs_06_22$service_date, format = "%V")
bs_06_22_week<- bs_06_22 %>% filter(week_num == "23" )

bs_07_22$week_num <- strftime(bs_07_22$service_date, format = "%V")
bs_07_22_week<- bs_07_22 %>% filter(week_num == "27" )

bs_08_22$week_num <- strftime(bs_08_22$service_date, format = "%V")
bs_08_22_week<- bs_08_22 %>% filter(week_num == "31" )

bs_09_22$week_num <- strftime(bs_09_22$service_date, format = "%V")
bs_09_22_week<- bs_09_22 %>% filter(week_num == "36" )

bs_10_22$week_num <- strftime(bs_10_22$service_date, format = "%V")
bs_10_22_week<- bs_10_22 %>% filter(week_num == "40" )

#combine the heavy rail data
bs_week<- rbind(bs_11_21_week, bs_12_21_week, bs_01_22_week, bs_02_22_week, bs_03_22_week, bs_04_22_week, bs_05_22_week, bs_06_22_week, bs_07_22_week, bs_08_22_week, bs_09_22_week, bs_10_22_week)
```



```{r}
hr_week %>% 
  ggplot(aes(travel_time_sec)) + 
  geom_density() +
  facet_wrap(week_num)
```

#Heavy rail and Light rail
```{r}
ggplot(hr_week) +
  aes(
    x = service_date,
    y = travel_time_sec,
    colour = route_id
  ) +
  geom_col() +
  scale_color_manual(
    values = c(Blue = "#4438F9",
    Orange = "#FF7B1C",
    Red = "#F6374D")
  ) +
  theme_minimal()

ggplot(lr_week) +
  aes(
    x = service_date,
    y = travel_time_sec,
    colour = route_id
  ) +
  geom_col() +
  scale_color_manual(
    values = c(Blue = "#4438F9",
    Orange = "#FF7B1C",
    Red = "#F6374D",
    `Green-B` = "#008900",
    `Green-C` = "#00C400",
    `Green-D` = "#00ED00",
    `Green-E` = "#45FF45",
    Mattapan = "#A7A7A7"
    )
  ) +
  theme_minimal()
```


#Comparison of scheduled time and actual time
```{r}
bs_03_22_week %>%
 filter(service_date %in% c("2022-03-07", "2022-03-13")) %>%
 ggplot() +
  aes(x = scheduled, y = actual, colour = service_date) +
  geom_tile(alpha= 0.3) +
  scale_color_manual(
    values = c(`2022-03-07` = "#1CA0FF",`2022-03-13` = "#FF4277")
  ) +
  theme(axis.text.x=element_blank(), 
        axis.text.y=element_blank(), 
        )
  theme_minimal()
```

```{r}
bs_03_22_week %>%
 filter(service_date %in% c("2022-03-07", "2022-03-13")) %>%
 filter(scheduled_headway >= 0L & scheduled_headway <= 6000L | is.na(scheduled_headway)) %>%
 filter(headway >= 0L & headway <= 
 6000L | is.na(headway)) %>%
 ggplot() +
  aes(
    x = scheduled_headway,
    y = headway,
    colour = service_date
  ) +
  geom_jitter(size = 1.5, alpha= 0.5) +
  geom_abline(a= 0,b= 1)
  scale_color_manual(
    values = c(`2022-03-07` = "#FF6156", `2022-03-13` = "#5F90FB")
  ) +
  theme_minimal()
```
This shows the difference of the scheduled time between the trip and the previous trip at the stop between Sunday and Monday.2022-03-07 is Monday and 2022-03-13 is Sunday. If the point is below the line, it means the time that bus stay at the stops is shorter than scheduled. Vice versa. From the graph we can see that those blue points are moving to the right overall compared to the red dots, which means that the headaway time of Sunday is overall longer than Monday. 





























